Parsing Postfix Log Files

### 2009-04-13 doc/logparser-complications.tex
Some minor changes.

    \item Is \daemon{smtpd} the only Postfix component that produced a log
        line for this mail?  Every mail that passes normally through
        Postfix will have a \daemon{cleanup} line, and later a
        \daemon{qmgr} log line; lack of a \daemon{cleanup} line is a sure
        sign the mail did not make it too far.

@@

During the \action{TIMEOUT} action, if the \daemon{cleanup} line is still
pending, the action updates the \texttt{timeout\_queueids} state table,
adding the queueid and the timestamp from the log line.  When the next
\daemon{cleanup} line is parsed for that queueid, the list will be checked
and if the log line meets the following requirements it will be deemed part
of the mail where the timeout occurred and discarded.

>>>>

    \item Is \daemon{smtpd} the only Postfix component that produced a log
        line for this mail?  Every mail that passes normally through
        Postfix will have a \daemon{cleanup} log line, and later a
        \daemon{qmgr} log line; lack of a \daemon{cleanup} log line is a
        sure sign the mail did not make it too far.

@@

During the \action{TIMEOUT} action, if the \daemon{cleanup} log line is
still pending, the action updates the \texttt{timeout\_queueids} state
table, adding the queueid and the timestamp from the log line.  When the
next \daemon{cleanup} log line is parsed for that queueid, the list will be
checked and if the log line meets the following requirements it will be
deemed part of the mail where the timeout occurred and discarded.

### 2009-04-13 doc/logparser-other-parsers-review.tex
Some minor changes.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing Splunk it took roughly 30 minutes to figure out how to
add a single log file to be indexed so that it could be searched, with the
downside that the log file was copied into a spool directory before
indexing, doubling the disk space usage.  The next test was to index all
the log files in a particular directory, but after three hours, numerous
futile attempts, and reading all the available documentation, the author
was still unable to index all the log files in a directory using the web
interface.  Using the \acronym{CLI} rather than the web interface was
partially successful: the command ``\texttt{splunk find logs
}\textit{log-directory\/}'' added 40 of the \numberOFlogFILES{} log files
to the queue for indexing.  Further attempts enqueued the same 40 log
files, without explaining why the others were excluded.\footnote{The log
files appear to have been indexed once only; presumably Splunk keeps track
of the log files it has indexed and discards requests to index log files
for a second time.  This may or may not be a useful feature for
\parsername{}.} There did not appear to be an option to ensure the log
files would be processed in the order they were created, though this may be
neither necessary nor beneficial with Splunk.  Subsequently the author was
successful in adding a single file at a time using the \acronym{CLI} and
the command ``\texttt{splunk add tail }\textit{filename\/}''.  A simple
loop using that command was enough to add all the desired log files.
Splunk will periodically check all indexed log files for updates unless
they are manually removed from its list; this may or may not be useful
behaviour.  Splunk did not appear to have any difficulty in indexing the
log files, once they had been successfully added to its queue.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\acronym{SMTP} or
\acronym{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \acronym{IP} address and hostname, queueid, start and end times,
timestamps for each log line, \acronym{SMTP} and enhanced status codes,
delivery delays, and message ID), though it does not make the full text of
the line available (this could be trivially added if desired, but would
greatly increase the size of the resulting database).  The full power of
\acronym{SQL} is available when searching the data extracted by
\parsername{}, allowing the user to search on arbitrarily complicated
conditions.

>>>>

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing Splunk it took roughly 30 minutes to figure out how to
add a single log file to be indexed so that it could be searched, with the
downside that the log file was copied into a spool directory before
indexing, doubling the disk space usage.  The next test was to index all
the log files in a particular directory, but after three hours, numerous
futile attempts, and reading all the available documentation, the author
was still unable to index all the log files in a directory using the web
interface.  Using the \acronym{CLI} rather than the web interface was
partially successful: the command ``\texttt{splunk find logs
}\textit{log-directory\/}'' added 40 of the \numberOFlogFILES{} log files
to the queue for indexing.  Further attempts enqueued the same 40 log
files, without explaining why the others were excluded.\footnote{The log
files appear to have been indexed once only; presumably Splunk keeps track
of the log files it has indexed and discards requests to index log files
for a second time.  This may or may not be a useful feature for
\parsername{}.} There did not appear to be an option to ensure the log
files would be processed in the order they were created, though this may be
neither necessary nor beneficial with Splunk.  Subsequently the author was
successful in adding a single log file at a time using the \acronym{CLI}
and the command ``\texttt{splunk add tail }\textit{filename\/}''.  A simple
loop using that command was enough to add all the desired log files.
Splunk will periodically check all indexed log files for updates unless
they are manually removed from its list; this may or may not be useful
behaviour.  Splunk did not appear to have any difficulty in indexing the
log files, once they had been successfully added to its queue.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\acronym{SMTP} or
\acronym{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \acronym{IP} address and hostname, queueid, start and end times,
timestamps for each log line, \acronym{SMTP} and enhanced status codes,
delivery delays, and message ID), though it does not make the full text of
the log line available (this could be trivially added if desired, but would
greatly increase the size of the resulting database).  The full power of
\acronym{SQL} is available when searching the data extracted by
\parsername{}, allowing the user to search on arbitrarily complicated
conditions.

### 2009-04-13 doc/logparser-results.tex
Some minor changes.

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 \& 62--68, despite their
larger than usual size.  Both groups of log files are much larger than
usual because of a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the mean number of rules tried per
log line reduces and so does the mean parsing time per log line.
\Graphref{Number of rules per Postfix component} shows the number of rules
per Postfix component, \graphref{Mean number of rules tried per log line}
shows the drop in the mean number of rules tried per log line for log files
containing a mail loop, and \graphref{Mean number of rules tried per log
line for each Postfix component} shows the mean number of rules tried per
log line for each Postfix component.

\showtable{build/include-file-size-and-number-of-log-lines-vs-parsing-time}{Ratio
of file size and number of log lines to parsing time}{parsing time vs file
size vs number of log lines factor table}

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of log lines for \numberOFlogFILES{} log files}{parsing time vs
file size vs number of log lines graph}

\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of file size
and number of log lines to parsing time (higher is better)}{parsing time vs
file size vs number of log lines factor}

>>>>

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs log file size vs number of log lines graph} shows
the parsing time in seconds, log file size in MB, and number of log lines
in tens of thousands, for each of the \numberOFlogFILES{} log files.  The
three lines run roughly in parallel, giving the impression that the
algorithm scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs log file size vs number of log lines factor},
which plots both the ratio of log file size vs parsing time, and the ratio
of number of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs log file size vs number of log lines factor
table} shows the same ratios for different groups of log files.  The ratios
are quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 \& 62--68, despite their
larger than usual size.  Both groups of log files are much larger than
usual because of a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the mean number of rules tried per
log line reduces and so does the mean parsing time per log line.
\Graphref{Number of rules per Postfix component} shows the number of rules
per Postfix component, \graphref{Mean number of rules tried per log line}
shows the drop in the mean number of rules tried per log line for log files
containing a mail loop, and \graphref{Mean number of rules tried per log
line for each Postfix component} shows the mean number of rules tried per
log line for each Postfix component.

\showtable{build/include-file-size-and-number-of-log-lines-vs-parsing-time}{Ratio
of log file size and number of log lines to parsing time}{parsing time vs
log file size vs number of log lines factor table}

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, log file
size, and number of log lines for \numberOFlogFILES{} log files}{parsing
time vs log file size vs number of log lines graph}

\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of log file
size and number of log lines to parsing time (higher is better)}{parsing
time vs log file size vs number of log lines factor}

### 2009-04-13 doc/logparser.tex
Minor changes.
Corrections to the first coverage section.
Renamed Results chapter to Evaluation. Finished evaluation chapter introduction and summary.
Ran diction and made changes based on its suggestions.
Made corrections based on Ariane's comments for chapters 1 and 2. Acronyms are now expanded the first time they're used in each chapter, rather than just the first time they're used in the thesis.
Some minor changes.

% Embed SVN Id etc.
\usepackage{svn}

@@

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

@@

@@

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.50\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.00\%%
}

@@

\title{Parsing Postfix log files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

@@

    XXX ASK CARL\@: SHOULD I HAVE A ``FUTURE WORK'' SECTION SOMEWHERE IN
    THE CONCLUSION\@?

@@

    XXX SPELL CHECK AND RUN DICTION\@.

    XXX MAKE IT EASY TO REDUCE THE HEIGHT OF GRAPHS\@.

@@

\input{logparser-introduction}
\input{logparser-background}
\input{logparser-other-parsers-review}
\input{logparser-architecture}
\input{logparser-implementation}
\input{logparser-results}
\input{logparser-conclusion}

>>>>

% Embed SVN Id
\usepackage{svn}

@@

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix Log Files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

@@

% This is silly, but it keeps chktex happy.
\newcommand{\singlequote}[0]{'}%

@@

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.500\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.000\%%
}

@@

\title{Parsing Postfix Log Files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

@@

@@

@@

\input{logparser-introduction}
% Reset the list of acronyms that have been used, so they will be expanded
% again the first time they are used.  The idea is to have them expanded
% once per chapter; I do not reset them before the appendices, but some may
% not have been used since they were reset, and thus will be expanded
% again.
\glsresetall{}
\input{logparser-background}
\glsresetall{}
\input{logparser-other-parsers-review}
\glsresetall{}
\input{logparser-architecture}
\glsresetall{}
\input{logparser-implementation}
\glsresetall{}
\input{logparser-results}
\glsresetall{}
\input{logparser-conclusion}
